---
title: "levante-r-tutorial"
format:
  html: default
date: "2026-07-01"
---

The levante R package documentation is available here: https://levante-framework.github.io/levante-r/index.html
This tutorial is adapted from a walkthrough available at the link above.

Access to the `levante-data-pilots-hackathon` dataset in this tutorial requires creating a Redivis (data-sharing platform) account, joining the LEVANTE organization, and completing a Data Use Agreement. For more information, see https://researcher.levante-network.org/data. 

## Setup 
```{r}
# install.packages("devtools") # if you need the devtools package, uncomment this line
devtools::install_github("levante-framework/levante-r")

# load levante and other supporting packages
install.packages("pacman")
library(pacman)
p_load(levante, dplyr, ggplot2)

# set a theme
theme_set(theme_classic())
```

## Using the `levante` package to load LEVANTE data
- Each table from Redivis can be read into R as a data frame using a specific `levante` `get_()` function.  
- Each `get_()` function takes at least two arguments, including a dataset name (required) and version number (defaults to the current version unless otherwise specified).  
- `get_participants()` is the first `levante` function in this code. When you run it, you'll be prompted to authenticate with a pop-up browser on Redivis. Subsequent `levante` calls to this dataset within the same session should not require additional authentication.  

```{r}
# Pull participant data from Redivis and save it to a data frame called "participants"
participants <- get_participants(data_source = "levante-data-pilots-hackathon:fpx0", version = "current") 

# What's in the participants data frame? 
head(participants)

# Now let's do some preliminary examination of the participants data
participants |> count(dataset, sort = TRUE) # the public data releases contain multiple datasets stapled together -- how many participants are there from each dataset?
```

Use `get_scores()` to obtain scored cognitive task data. 
```{r}

# Pull scored data from Redivis and save it to a data frame called "scores"
scores <- # complete this line of code to get scores from levante-data-pilots-hackathon
  
# What's in the scores data frame?
head(scores)

# Now let's do some preliminary examination of the scores data
scores |> count() # complete this line of code code to return the number of scores for each task_id 
```

```{r}
# This plot returns scores by age for each task.
ggplot(scores, aes(x = age, y = score)) +
  facet_wrap(vars(task_id)) +
  geom_point()
```

Use `get_trials()` to access trial-level data. Note that this is large and may take longer to load.
```{r}

# Pull trial-level data from Redivis and save it to a data frame called "trials"
trials <- # complete this line of code to get scores from levante-data-pilots-hackathon

# What's in the trials data frame?
head(trials)

# Now let's do some preliminary checks on the trials data
trials |> count() # complete this line of code to return the number of trials per dataset

trials |> count() # complete this line of code to return the number of trials per task_d
```

Use `get_surveys()` to access item-level survey data.
```{r}

# Pull survey data from Redivis and save it to a data frame called "surveys"
surveys <- # complete this line of code to get scores from levante-data-pilots-hackathon

# What's in the surveys data frame?
head(trials)

# Now let's do some preliminary examination of the trials data
surveys |> count() # complete this line of code to return the number of surveys per dataset
```

Use `get_parameters()` to access the IRT item parameters used in LEVANTE scoring.
```{r}

# Pull up-to-date item parameters from Redivis
item_parameters <-# complete this line of code to get scores from levante-data-pilots-hackathon

item_parameters |> count() # returns the number of parameters per task_id
```

Researchers accessing their own LEVANTE data can use `get_raw_data()` to access additional tables within private dataset. `get_raw_data()` is a general function that can pull any table. This function is not covered as a part of this tutorial, which focuses on accessing the public dataset (see documentation linked at the top of this code for more info).