---
title: "Multigroup IRT in LEVANTE tutorial"
author: "Mika Braginsky"
---

This tutorial shows you the basics of how to fit and use Item Response Theory models using the `mirt` R package.
We'll use a simplified version the trial-level data from the LEVANTE mental rotation task taken from the `levante-data-pilots-hackathon` dataset -- subset to one response per participant per trial to avoid complication in the data preparation and model set up.

Places marked `TODO` are where you're expected to fill in code.

```{r}
library(mirt)
library(tidyverse)
```


# Step 1: Load in data and prep it for modeling

```{r}
# read in simplified mental rotation trial-level data
trials <- read_rds("data/trials_mrot.rds")

# get a feel for sample sizes
trials |> distinct(dataset, run_id) |> count(dataset) # how many runs in each dataset
trials |> count(item_uid) # how many responses for each item
```

```{r}
trials_wide <- trials |>
  # convert values to numeric
  mutate(value = as.numeric(correct)) |>
  # subset columns to run, dataset, item, value
  select(run_id, dataset, item_uid, value) |>
  # change structure to a column for each item
  pivot_wider(names_from = item_uid, values_from = value) |>
  # move run_id to rownames
  column_to_rownames("run_id")

# split up wide data into run x item matrix and dataset vector
trials_prepped <- trials_wide |> select(-dataset)
run_groups <- trials_wide |> pull(dataset)
```


# Step 2: Fit models

```{r}
# fit SingleGroup Rasch model
m_pooled_rasch <- mirt(
  data = trials_prepped,
  itemtype = "Rasch",
  guess = 0.5
)

# fit SingleGroup 2PL model
m_pooled_2pl <- mirt(
  # TODO: copy arguments from above, change itemtype to 2PL
)
```

```{r}
# define invariance arguments
invariances <- list(
  configural = "",
  metric = c("slopes"),
  scalar = c("free_means", "free_var", "intercepts", "slopes")
)

# fit MultipleGroup Rasch model with configural invariance
m_configural_rasch <- multipleGroup(
  data = trials_prepped,
  group = run_groups,
  itemtype = "Rasch",
  invariance = invariances$configural,
  guess = 0.5
)

# fit MultipleGroup Rasch model with scalar invariance
m_scalar_rasch <- multipleGroup(
  # TODO: change invariance to scalar
)

# fit MultipleGroup 2PL model with configural invariance
m_configural_2pl <- multipleGroup(
  # TODO: change itemtype to 2PL (invariance back to configural)
)

# fit MultipleGroup 2PL model with metric invariance
m_metric_2pl <- multipleGroup(
  # TODO: change invariance to metric
)

# fit MultipleGroup 2PL model with scalar invariance
m_scalar_2pl <- multipleGroup(
  # TODO: change invariance to scalar
)
```

# Step 3: Extract various outputs from models

## Model-level statistics

```{r}
# which models fit better according to BIC?
anova(m_scalar_rasch, m_scalar_2pl, m_configural_rasch, m_metric_2pl, m_configural_2pl)

# various properties of model objects can be extracted (see ?extract.mirt)
extract.mirt(m_scalar_2pl, "BIC")    # BIC
extract.mirt(m_scalar_2pl, "nest")   # number of estimated parameters

# marginal reliability of SingleGroup models
marginal_rxx(m_pooled_rasch)
marginal_rxx(m_pooled_2pl)

# marginal reliability of MulipleGroup models, per group
m_scalar_2pl_de <- extract.group(m_scalar_2pl, "pilot_mpieva_de_main")
marginal_rxx(m_scalar_2pl_de)
# can also see what the reliability is depending on ability
plot(m_scalar_2pl_de, type = "rxx")

# iterating over all groups
unique(run_groups) |> set_names() |>
  map(\(gr) extract.group(m_scalar_2pl, gr) |> marginal_rxx())
```

## Item-level parameters

```{r}
# item parameters
coef(m_configural_2pl, simplify = TRUE)
coef(m_scalar_2pl, simplify = TRUE)

# item response functions
plot(m_configural_2pl, type = "trace")
plot(m_scalar_2pl, type = "trace")
```

## Assessment/person-level scores

```{r}
scores <- fscores(m_scalar_2pl, full.scores.SE = TRUE) |>
  as_tibble() |>
  rename(ability = "F1", se = "SE_F1") |>
  mutate(run_id = rownames(trials_prepped), .before = everything())

hist(scores$ability, breaks = 20) # distribution of scores
hist(scores$se, breaks = 20)      # distribution of SEs
```


# Exercise

Which type of model would have more estimated parameters, a Rasch model with configural invariance or a Rasch model with scalar invariance? How about a Rasch model with scalar invariance or a 2PL model with scalar invariance? Make a guess based on your understanding, and then check it using `extract.mirt()`.

```{r}
extract.mirt() # TODO
```


# Resources

See also documentation for `mirt`:
https://philchalmers.github.io/mirt

LEVANTE model fitting code is in the `levantemodels` R package:
https://levante-framework.github.io/levantemodels

Fitted models and scoring specifications are saved in the levante-metadata-scoring dataset:
https://stanford.redivis.com/datasets/e97h-009zz0c49

Note some complications in our real data/models:
- participants can see given item multiple times (actually same item or collapsed item types)
- set of presented items can be different across sites
- items can have all correct/incorrect responses
- include priors over parameter values
- functionality for iteration over tasks/languages
- models are converted to model record objects and saved in model registry
- post hoc scoring from pre-fit models (also with potential item alignment issues)
